---
title: "p8105_hw2_rc3520.Rmd"
author: "Ruyi Chen"
date: "2023-10-03"
output: github_document
---

```{r}
library (tidyverse)
library(tidyr)
library(readxl)
library(readr)
```


# Probelm 1
## Clean the data in pols-month.csv.
### Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.
```{r eval=FALSE}
 pols_month_df= 
  read_csv("./data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(col = mon,into = c("year","month","day"),sep = '-', convert = TRUE) %>%
  mutate(
    year = as.integer(year), 
    month = as.integer(month), 
    month = month.abb[month],
    day = as.integer(day),
    president = ifelse(prez_gop, "dem", "gop")
    ) %>%
  select(-prez_dem, -prez_gop, -day)
pols_month_df
```

## Clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.
```{r eval = False}
snp_df=
    read_csv("./data/snp.csv") %>%
    drop_na() %>%
    janitor::clean_names() %>%
    separate(date, into  = c("month", "day", "year")) %>%
  mutate(year = as.integer(year), 
         year=ifelse(year<23, year+2000, year+1900),
         month = as.integer(month), 
         month = month.abb[month],
         day = as.integer(day)) %>%
  select(year,month,close,-day) %>%
  arrange(year,month) 
snp_df

```
##Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.
```{r eval = False}
unemployment_df=
    read_csv("./data/unemployment.csv",
             col_names = TRUE,
              show_col_types = FALSE) %>%
    janitor::clean_names() %>%
  drop_na() %>%
  pivot_longer(jan:dec, names_to = "month", values_to = "unemployment_percent") %>%
  mutate(month = str_to_title(month)) %>%
  arrange(year, month)
unemployment_df
```
##Join the datasets by merging snp into pols, and merging unemployment into the result
```{r eval = False}
pols_snp_df = left_join(pols_month_df, snp_df, by = c("year", "month"))
pols_snp_unemployment_df = left_join(pols_snp_df, unemployment_df, by = c("year", "month"))

pols_snp_unemployment_df

```
##Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).
pols_month_df dataset includes 822 observations and 9 variables (key variables include year, month and president) containing information about the national politicians 
These data describe whether the policians are democratic or republican between 1947 and 2015. It also contains  the number of national politicians(governors/senators/representatives) in U.S. I create a new variable president in order to combine pre_gop and prez_dem.
snp_df includes 787 observations and 3 variables including year, month and close. They describe the closing values of S&P stock index from between 1950 and 2015.
unemployment_df includes 804 observations and 3 variables including year, month and unemployment_rate.  It shows the unemployment rate in each month between 1948 and 2014.
pols_snp_unemployment_df shows the number of republic / democratic presidents and number of governors/ senators/ representatives from 1947 to 2015. It also shows the closing values of the S&P stock index between 1950 and 2015 and unemployment rate between Jan/1948 to Jun/2014.

#Problem 2
##clean the Mr. Trash Wheel sheet
```{r}
Mr_trash_wheel = read_excel("./data/202309 Trash Wheel Collection Data.xlsx",
                     sheet = "Mr. Trash Wheel",
                     range=cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(
    sports_balls = as.integer(round(sports_balls)), 
    year = as.numeric(year), 
    dumpster = as.numeric(dumpster), 
    homes_powered = as.integer(round(homes_powered, digits = 0))) 

Mr_data
```



