---
title: "p8105_hw2_rc3520.Rmd"
author: "Ruyi Chen"
date: "2023-10-03"
output: github_document
---

```{r}
library (tidyverse)
library(tidyr)
library(readxl)
library(readr)
```


# Probelm 1
## Clean the data in pols-month.csv.
### Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.
```{r eval=FALSE}
 pols_month_df= 
  read_csv("./data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(col = mon,into = c("year","month","day"),sep = '-', convert = TRUE) %>%
  mutate(
    year = as.integer(year), 
    month = as.integer(month), 
    month = month.abb[month],
    day = as.integer(day),
    president = ifelse(prez_gop, "dem", "gop")
    ) %>%
  select(-prez_dem, -prez_gop, -day)
pols_month_df
```

## Clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.
```{r eval=FALSE}
snp_df=
    read_csv("./data/snp.csv") %>%
    drop_na() %>%
    janitor::clean_names() %>%
    separate(date, into  = c("month", "day", "year")) %>%
  mutate(year = as.integer(year), 
         year=ifelse(year<23, year+2000, year+1900),
         month = as.integer(month), 
         month = month.abb[month],
         day = as.integer(day)) %>%
  select(year,month,close,-day) %>%
  arrange(year,month) 
snp_df

```
##Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.
```{r eval=FALSE}
unemployment_df=
    read_csv("./data/unemployment.csv",
             col_names = TRUE,
              show_col_types = FALSE) %>%
    janitor::clean_names() %>%
  drop_na() %>%
  pivot_longer(jan:dec, names_to = "month", values_to = "unemployment_percent") %>%
  mutate(month = str_to_title(month)) %>%
  arrange(year, month)
unemployment_df
```
##Join the datasets by merging snp into pols, and merging unemployment into the result
```{r eval = FALSE}
pols_snp_df = left_join(pols_month_df, snp_df, by = c("year", "month"))
pols_snp_unemployment_df = left_join(pols_snp_df, unemployment_df, by = c("year", "month"))

pols_snp_unemployment_df

```
##Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).
pols_month_df dataset includes 822 observations and 9 variables (key variables include year, month and president) containing information about the national politicians 
These data describe whether the policians are democratic or republican between 1947 and 2015. It also contains  the number of national politicians(governors/senators/representatives) in U.S. I create a new variable president in order to combine pre_gop and prez_dem.
snp_df includes 787 observations and 3 variables including year, month and close. They describe the closing values of S&P stock index from between 1950 and 2015.
unemployment_df includes 804 observations and 3 variables including year, month and unemployment_rate.  It shows the unemployment rate in each month between 1948 and 2014.
pols_snp_unemployment_df shows the number of republic / democratic presidents and number of governors/ senators/ representatives from 1947 to 2015. It also shows the closing values of the S&P stock index between 1950 and 2015 and unemployment rate between Jan/1948 to Jun/2014.

#Problem 2
##clean the Mr. Trash Wheel sheet
```{r}
mr_trash_wheel = read_excel("./data/202309 Trash Wheel Collection Data.xlsx",
                     sheet = "Mr. Trash Wheel",
                     range=cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  drop_na() %>%
  mutate(
    sports_balls = as.integer(round(sports_balls)), 
    year = as.integer(year), 
    dumpster = as.numeric(dumpster), 
    homes_powered = (weight_tons * (500/30)),
    catg = "mr trash wheel") %>%
    # Create a variable catg to help with the combination of the datasets
  select(catg, everything())
mr_trash_wheel
```

##import, clean, and organize the data for Professor Trash Wheel
```{r}
professor_trash_wheel = read_excel("./data/202309 Trash Wheel Collection Data.xlsx",
                                   sheet = "Professor Trash Wheel")%>%
                        janitor::clean_names() %>%
                        drop_na(dumpster) %>%
                        mutate(homes_powered = (weight_tons * (500/30)),
                          year=as.integer(year),
                          catg = "prof trash wheel") %>%
  # Create a variable catg to help with the combination of the datasets
  select(catg, everything())
professor_trash_wheel
```
## calculate total weight of trash collected by Professor Trash Wheel
```{r}
professor_trash_wheel %>%
  select(weight_tons) %>%
  colSums()
```

##import, clean, and organize the data for Gwynnda
```{r}
gwynnda = read_excel("./data/202309 Trash Wheel Collection Data.xlsx",
                                   sheet = "Gwynnda Trash Wheel")%>%
                        janitor::clean_names() %>%
                        drop_na(dumpster) %>%
                        mutate(homes_powered = (weight_tons * (500/30)),
                          cigarette_butts = as.integer(cigarette_butts),
                          year=as.integer(year),
                          catg = "gwynnda trash wheel") %>%
    # Create a variable catg to help with the combination of the datasets
  select(catg, everything())
gwynnda
```

##Calculate The total number of cigarette butts collected by Gwynnda in July of 2021
```{r}
gwynnda %>%
  filter(year == "2021",month == "July")%>%
  select(cigarette_butts)%>%
  colSums()
```

##Combine these there data to produce a single tidy dataset

```{r}
mr_prof_trash_wheel = left_join(mr_trash_wheel, professor_trash_wheel, by = c("dumpster"))
mr_prof_gwynnda_trash_wheel= left_join(mr_prof_trash_wheel,gwynnda , by = c("dumpster"))
mr_prof_gwynnda_trash_wheel
```

##Description about the datasets
mr_trash_wheel includes `r nrow(mr_trash_wheel)` observations and `r ncol(mr_trash_wheel)`variables.Key variables include year, month, dumpster and weight.
professor_trash_wheel includes `r nrow(professor_trash_wheel)`  observations and  `r ncol(professor_trash_wheel)` variables. Key variables include year, month, dumpster and weight.
The total weight of trash collected by Professor Trash Wheel is `r professor_trash_wheel %>%
  select(weight_tons) %>%
  colSums()` tons.
gwynnda includes  `r nrow(gwynnda)` observations and `r ncol(gwynnda)` variables. Key variables include year, month, dumpster, weight, and cigarette butts.
The total number of cigarette butts collected by Gwynnda in July of 2021 is `r gwynnda %>%
  filter(year == "2021",month == "July")%>%
  select(cigarette_butts)%>%
  colSums()`
  
# Problem 3
##Import, clean, and tidy the dataset of mci_baseline 
```{r}
mci_baseline =
  read_csv("./data/MCI_baseline.csv", skip = 1) %>% 
  janitor::clean_names() %>%
  mutate(
    sex = case_match(sex, 1 ~ "Male",0 ~ "Female"),
    apoe4 = case_match(apoe4, 1 ~ "Carrier", 0 ~ "Non_carrier")) 

mci_baseline_df = filter(mci_baseline, age_at_onset == "." | age_at_onset > current_age)
mci_baseline_df
```
##Discussion on the dataframe mci_baseline
After importing the data, I tidy the variable names by using 'janitor::clean_names' and deleting the first line(description on the variables) by using 'skip'.Then, I make sure that sex and APOE4 carrier status are appropriately encoded by using 'mutate' so that 1 and 0 will be "Male" and "Female" under sex and will be "Carrier" and "Non_carrier" under apoes. And I also use 'filter' and generate the dataframe 'mci_baseline_df' to represent the data that meets the stated inclusion criteria, either missing `age_at_onset` value or `age_at_onset` is larger than `current_age` at baseline.
### `r nrow(mci_baseline)` numbers of patients were recruited.
### `Of these `r nrow(mci_baseline_df)` developed MCI.
### Average baseline age is `r mean(as.numeric(filter(mci_baseline_df, age_at_onset != ".")$current_age))`.
### `r nrow(filter(mci_baseline_df, sex=="Female"& apoe4=="Carrier")) / nrow(filter(mci_baseline, sex== "Female")) * 100` % of women in the study are APOE4 carriers.


##Import, clean, and tidy the dataset of mci_amyloid 
```{r}
mci_amyloid =
  read_csv("./data/mci_amyloid.csv", skip = 1) %>% 
  janitor::clean_names() %>%
  rename(id = study_id)
mci_amyloid
```
##Description:
After importing the data with read_csv, I tidy the variable names by using janitor::clean_names and deleting the first line(description on the variables) by using skip. I also use rename to change the column name "study_id" to "id" so that it can be further match to the column name in mci_baseline.There are `r nrow(mci_amyloid)` people enrolled in this study. There are `r ncol(mci_amyloid)` vairables including id and biomarker measurement recorded at baseline,time_2, time_4, time_6 and time_8.

##Combine the demographic and biomarker datasets 
```{r}
mci_baseline_amyloid = 
  inner_join(mci_baseline_df, mci_amyloid) 
#export the result as a CSV to your data directory.
write_csv(mci_baseline_amyloid, "./data/baseline_amyloid.csv")
```
##Description:
By applying inner join, I keep participants who exist on both demographic and biomarker datasets. There are `r nrow(mci_baseline_amyloid)` numbers of participants who appear on both datasets. Since there are `r nrow(mci_amyloid)-nrow(mci_baseline_df)` more participants in biomarker datasets than participants in demographic datasets, some participants appear in only the baseline or amyloid datasets. Among these `r nrow(mci_baseline_amyloid)` participants, `r nrow(filter(mci_baseline_amyloid, apoe4=="Carrier")) ` are carriers which 
`r nrow(filter(mci_baseline_amyloid, sex=="Female"& apoe4=="Carrier")) / nrow(filter(mci_baseline_amyloid, apoe4=="Carrier")) * 100` % of them are females and `r nrow(filter(mci_baseline_amyloid, sex=="Male"& apoe4=="Carrier")) / nrow(filter(mci_baseline_amyloid, apoe4=="Carrier")) * 100` % of them are males; `r nrow(mci_baseline_amyloid)` participants, `r nrow(filter(mci_baseline_amyloid, apoe4=="Non_carrier")) ` are non_carriers which `r nrow(filter(mci_baseline_amyloid, sex=="Female"& apoe4=="Non_carrier")) / nrow(filter(mci_baseline_amyloid, apoe4=="Non_carrier")) * 100` % of them are females and `r nrow(filter(mci_baseline_amyloid, sex=="Male"& apoe4=="Non_carrier")) / nrow(filter(mci_baseline_amyloid, apoe4=="Non_carrier")) * 100` % of them are males.

